On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-01-21 MS.tex
Small editing in ms

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across stages. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively. 
\item Despite a sizeable literature on traits and growth, major knowledge gaps across size-classes still exist; for one-third of traits the majority of correlations with growth ($>$ 75\%) come from a single and unique size-classe. Clearly, there is scope for strengthening our understanding on the link between traits and growth through adding further empirical studies for traits and size-classes underrepresented in the literature.

@@

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes (i.e. seedling, sapling, and adult). We therefore used plant stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms. 
We chose to assess our hypotheses via a meta-analysis because there already exist a considerable number of studies with salient information, allowing us to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between trait and growth.

@@

The rate at which \textbf{live biomass} $B$ is produced by a plant is given by:
\begin{equation}\label{eq:eq0}
\underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}} = \underbrace{\alpha_{\rm y}}_{\text{yield}}  \big(\underbrace{A_{\rm l} p}_{\text{photosynthesis}} - \underbrace{\sum_{i = {\rm l,s,r}} M_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i = \rm{l,s,r}} M_{i} k_{i}}_{\text{turnover}},
\end{equation}
where $\alpha_{\rm y}$ is the ratio of carbon fixed in mass per carbon assimilated, and $p$ is the assimilation rate of CO$_{2}$ per leaf area. Here, and in the following equations, $M_i$ denotes the \textbf{live mass on the plant}, $r_i$ the mass-specific \textbf{respiration rate}, and $k_i$ the \textbf{turnover rate} of different plant tissues; the subscripts $i = \rm{l,s,r,t,f}$ referring to leaves, sapwood and bark, roots, total of these vegetative tissues, and to the reproductive tissues (fruits, flowers etc.)  Similarly $A_i$ denotes \textbf{areas}, of leaves ($i = l$) and of cross-sections of total stem and sapwood ($i =\rm{st,ss}$) respectively. Note that the live biomass $B = M_{\rm t} + M_{\rm f}$.  Applying the product rule of differential calculus, the \textbf{growth rate in plant mass}, ($M_{\rm t}$), can be decomposed as
\begin{equation}\label{eq:eq1}
\underbrace{{\frac{\ud M_{\rm t}}{\ud t}}}_{\text{growth rate in mass}} = \underbrace{\frac{\ud M_{\rm t}}{\ud B}}_{\text{allocation to vegetative growth}}  \times \underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}},
\end{equation}
where $\frac{\ud M_{\rm t}}{\ud B}$ is the fraction of net live biomass increase that goes to vegetative growth rather than to production of reproductive tissues.

The \textbf{growth rate in height} can likewise be decomposed as the product of several meaningful terms:
\begin{equation}\label{eq:eq2}
\underbrace{\frac{\ud H}{\ud t}}_{\text{growth rate in height}} = \underbrace{\frac{\ud H}{\ud A_{\rm l}}}_{\text{architecture layout}} \times \underbrace{{\frac{\ud A_{\rm l}}{\ud M_{\rm t}}}}_{\text{leaf deployment per mass}} \times {\frac{\ud M_{\rm t}}{\ud B}} \times {\frac{\ud B}{\ud t}}.
\end{equation}
The first term on the right of eq \ref{eq:eq2},
$\ud H / \ud A_{\rm l}$, is the growth in plant height
per unit growth in total leaf area -- accounting for the architectural
strategy of the plant. We assume this function is unrelated to the traits considered in the current paper. The second term, $\ud A_{\rm l} / \ud M_{\rm t}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, $\ud A_{\rm l} / \ud M_{\rm t}$
can itself be expressed as a sum of construction costs per unit leaf
area,
\begin{equation}\label{eq:daldmt}
\frac{\ud A_{\rm l}}{\ud M_{\rm t}}
= \left(\frac{\ud M_{\rm l}}{\ud A_{\rm l}}
 +  \frac{\ud M_{\rm s}}{\ud A_{\rm l}} + \frac{\ud M_{\rm r}}{\ud A_{\rm l}}\right)^{-1}.
\end{equation}

@@

Plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ and $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ in the above model. Whether plant growth rate increases with SLA depends on the how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA also change with size (Table \ref{tab:trade-offs}). During early life, when leaf comprises a large part of the plant, decreasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ has a large effect compared to the other terms in eq. \ref{eq:daldmt}. As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the relative effect of leaf turnover comes to dominate at larger sizes, and as such, the effect of SLA on height and diameter growth shifts from positive to negative as plants increase in size.

Among seedlings, growth is often measured in units of plant mass, rather than stem or diameter. For mass growth in seedlings, cheaper leaf construction (high SLA)  translates directly into faster relative growth rate (See Appendix XXX eq. \ref{eq:eq1_seedlings_RGR}).

@@

High WD increases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table \ref{tab:trade-offs}), which should slow growth. This effects occur within the term $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ in the above model. Although expensive wood decreases the risk of mortality, this benefit is not reflected in growth outcomes. Thus increasing WD is predicted to decrease growth across the entire size range.

Cheaper stem construction (low WD or stem density) increases the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$), and consequently increases growth rate in leaf area, height, and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all plant stages. Because the relative expenditure on stem compared to leaf is increases as plants grow larger, the relationship between WD and growth may also strengthen in larger plants.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm t}}{\ud B}$ in the above model. For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth (i.e. $\frac{\ud M_{\rm t}}{\ud B} =1$). But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates (Table \ref{tab:trade-offs}).

@@

The main effect of increased seed size is to produce a larger overall seedling. Under similar light conditions, RGR decreases with size because the ratio of leaf area to total biomass decreases with increasing plant size. Thus, larger seedlings are predicted to have a lower RGR during the earliest phases of growth \citep{Turnbull:2012ew}. However, among larger plants we expect the effect of this initial size to be minimal, and thus the correlation between seed size and growth rate should disappear. The ${\rm RGR}_{{\rm mass}}$ slows as plants grow larger, and large-seeded species begin larger as seedlings. For larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}). However, the predicted relationship between growth and seed mass in seedlings reverses if AGR$_{\rm mass}$ is used instead of RGR$_{\rm mass}$.

@@

\subsection*{Size-classes representativity among 19 traits recorded}
In total, correlations had been reported for at least two different stages across 18 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits correlations would be reported only for one size class. But this limitation applied only for LAR and NARmass, where studies appeared confined to seedling stages. However, for a majority of traits the number of observations across stage was higthly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from a single and unique plant stage (Fig. \ref{fig:fig1}).

@@

TODO: XXXX Something about limited amount of data (add the following arguments:)
Despite a sizeable literature on traits and growth, there are major knowledge gaps on the link between trait and growth. The number of correlations reported for each 19 traits is generally unbalanced in favor of a particular stage, either for historical or practical reasons. This means that the knowledge we have on each trait is often biased toward a particular size-classe. Additional data are specifically required in small plants (seedling and sapling) for WD and Hmax, and in large plants (adult and sapling) for Aarea and more broadly for all traits requiring a measure of whole plant biomass (e.g. LMR, LAR).

@@

Clearly there is much scope for strengthening our understanding of correlations between traits and growth through (1) adding further empirical studies for traits and size-classes with weaker representation, (2) accounting more satisfactorily for variables such as light, and (3) expressing plant sizes on continuous scales of mass or height rather than through the approximate and partly subjective categories unavoidably adopted here. However, the existing published evidence is consistent with qualitative predictions of a simple model that assesses the effects of plant size on growth. Thus, the linkage of growth to the mechanisms outlined here can shed light on tree growth trajectories and the size-dependent effects of functional traits on the changing species growth rate rankings as individuals grow larger.

>>>>

\item Results were consistent with predictions, even with some limitations arising from unequal numbers of observation across stages. Specific leaf area was correlated with relative growth rate (RGR) in seedlings but not in adult plants. Correlations of growth with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with RGR only in one plant stage category: seedlings and adults, respectively. 
\item Despite a sizeable literature on traits and growth, major knowledge gaps across size still exist; for one-third of traits the majority ($>$ 75\%) of correlations with growth come from a single stage. Clearly, there is scope for strengthening our understanding on the link between traits and growth through adding further empirical studies for traits and size classes underrepresented in the literature.

@@

In this paper, we investigate for five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass):
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from diverse studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes (i.e. seedling, sapling, and adult). We therefore used plant stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms. 
We chose to assess our hypotheses via a meta-analysis because there already exist a considerable number of studies with salient information. Using these studies, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. While, we anticipated a reasonably-sizeable literature on trait and growth to be available, we were uncertain about the representation of all trait-stages combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between trait and growth.

@@

The rate at which \textbf{live biomass} $B$ is produced by a plant is given by
\begin{equation}\label{eq:eq0}
\underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}} = \underbrace{\alpha_{\rm y}}_{\text{yield}}  \big(\underbrace{A_{\rm l} p}_{\text{photosynthesis}} - \underbrace{\sum_{i = {\rm l,s,r}} M_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i = \rm{l,s,r}} M_{i} k_{i}}_{\text{turnover}} \, ,
\end{equation}
where $\alpha_{\rm y}$ is the ratio of carbon fixed in mass per carbon assimilated, and $p$ is the assimilation rate of CO$_{2}$ per leaf area. Here, and in the following equations, $M_i$ denotes the \textbf{live mass on the plant}, $r_i$ the mass-specific \textbf{respiration rate}, and $k_i$ the \textbf{turnover rate} of different plant tissues; the subscripts $i = \rm{l,s,r,t,f}$ referring to leaves, sapwood and bark, roots, total of these vegetative tissues, and to the reproductive tissues (fruits, flowers etc.)  Similarly $A_i$ denotes \textbf{areas}, of leaves ($i = l$) and of cross-sections of total stem and sapwood ($i =\rm{st,ss}$) respectively. Note that the live biomass $B = M_{\rm t} + M_{\rm f}$.  Applying the product rule of differential calculus, the \textbf{growth rate in plant mass}, ($M_{\rm t}$), can be decomposed as
\begin{equation}\label{eq:eq1}
\underbrace{{\frac{\ud M_{\rm t}}{\ud t}}}_{\text{growth rate in mass}} = \underbrace{\frac{\ud M_{\rm t}}{\ud B}}_{\text{allocation to vegetative growth}}  \times \underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}},
\end{equation}
where $\frac{\ud M_{\rm t}}{\ud B}$ is the fraction of net live biomass increase that goes to vegetative growth rather than to production of reproductive tissues.

The \textbf{growth rate in height} can likewise be decomposed as the product of several meaningful terms:
\begin{equation}\label{eq:eq2}
\underbrace{\frac{\ud H}{\ud t}}_{\text{growth rate in height}} = \underbrace{\frac{\ud H}{\ud A_{\rm l}}}_{\text{architecture layout}} \times \underbrace{{\frac{\ud A_{\rm l}}{\ud M_{\rm t}}}}_{\text{leaf deployment per mass}} \times {\frac{\ud M_{\rm t}}{\ud B}} \times {\frac{\ud B}{\ud t}}.
\end{equation}
where $\ud H / \ud A_{\rm l}$, is the growth in plant height
per unit growth in total leaf area -- accounting for the architectural
strategy of the plant. We assume this function is unrelated to the traits considered in the current paper. The second term, $\ud A_{\rm l} / \ud M_{\rm t}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, $\ud A_{\rm l} / \ud M_{\rm t}$
can itself be expressed as a sum of construction costs per unit leaf
area,
\begin{equation}\label{eq:daldmt}
\frac{\ud A_{\rm l}}{\ud M_{\rm t}}
= \left(\frac{\ud M_{\rm l}}{\ud A_{\rm l}}
 +  \frac{\ud M_{\rm s}}{\ud A_{\rm l}} + \frac{\ud M_{\rm r}}{\ud A_{\rm l}}\right)^{-1}.
\end{equation}

@@

Plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ and $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ in the above model. Whether plant growth rate increases with SLA depends on how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA change with size (Table \ref{tab:trade-offs}). 
During early life, when leaf comprises a large part of the plant, decreasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ has a large effect compared to the other terms in eq. \ref{eq:daldmt}. For mass growth, cheaper leaf construction (high SLA) translates directly into faster relative growth rate in seedlings (See Appendix XXX eq. \ref{eq:eq1_seedlings_RGR}). As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the relative effect of leaf turnover comes to dominate at larger sizes, and as such, the effect of SLA on height and diameter growth shifts from positive to negative as plants increase in size.

@@

Cheaper stem construction (low WD or stem density) increases the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm t}}$), and consequently increases growth rate in leaf area, height, and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all plant stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between WD and growth may also strengthen in larger plants.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm t}}{\ud B}$ in the above model eqs. (\ref{eq:eq2}-\ref{eq:eq5}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth (i.e. $\frac{\ud M_{\rm t}}{\ud B} =1$). But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates (Table \ref{tab:trade-offs}).

@@

The main effect of increased seed size is to produce a larger overall seedling. Under similar light conditions, RGR decreases with size because the ratio of leaf area to total biomass decreases along plant size. Thus, larger seedlings are predicted to have a lower RGR during the earliest phases of growth \citep{Turnbull:2012ew}. However, as plants are growing we expect the effect of this initial size to be minimal, and the correlation between seed size and growth rate should disappear. Thus for larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}). However, the predicted relationship between growth and seed mass in seedlings reverses if AGR$_{\rm mass}$ is used instead of RGR$_{\rm mass}$.

@@

\subsection*{Representation of stages among 19 traits}
In total, correlations had been reported for at least two different stages across 18 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits correlations would be reported only for one size class. But this limitation applied only for LAR and NARmass, where studies appeared confined to seedling stages. However, for a majority of traits the number of observations across stage was higthly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from a single stage (Fig. \ref{fig:fig1}).

@@

TODO: XXXX Something about limited amount of data (add the following arguments:)
Despite a sizeable literature on traits and growth, there are major knowledge gaps on the link between trait and growth. The number of correlations reported for each 19 traits is generally unbalanced in favor of a particular stage, either for historical or practical reasons. This means that the knowledge we have on each trait is often biased toward a particular size-classe.

@@

Clearly there is much scope for strengthening our understanding of correlations between traits and growth through (1) adding further empirical studies for traits and size-classes with weaker representation, (2) accounting more satisfactorily for variables such as light, and (3) expressing plant sizes on continuous scales of mass or height rather than through the approximate and partly subjective categories unavoidably adopted here. Additional data are specifically required in small plants (seedling and sapling) for WD and Hmax, and in large plants (adult and sapling) for Aarea and more broadly for all traits requiring a measure of whole plant biomass (e.g. LMR, LAR). However, the existing published evidence is consistent with qualitative predictions of a simple model that assesses the effects of plant size on growth. Thus, the linkage of growth to the mechanisms outlined here can shed light on tree growth trajectories and the size-dependent effects of functional traits on the changing species growth rate rankings as individuals grow larger.
